A survey of HTTP caching options on the open Web

### 2015-01-08 sanity/paper1.tex
write about the breakup
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@944 43ef1190-90fd-4dd7-a769-d0121f02b80e

@@

We next examine the freshness lifetime of the 2965 information
resources that returned a successful response, either with SPARQL
results or RDF data. 405 returned valid headers, but 114 did so to
\emph{prohibit} caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. Additionally, 269 resources had a \httph{Cache-Control}
header to control other aspects of caching than lifetime, i.e. to say
that only private caches may use the response, that the cache must be
revalidated, or to prohibit caching.

Caching may prohibited by setting a non-positive freshness
lifetime. Common practice is to set the \httph{Expires} header to a
date far back in time, which has also been done in large parts of the
present dataset. For the following, we have set any non-positive
lifetime to 0, as that is semantically equivalent. With that, we
obtain the summary statistics in Table~\ref{tab:summaryhard}. In
figure\todo{add figure}, we have included a histogram with the counts
of the resources that had 0 lifetime, and then boxes of varying length.

@@

@@

\begin{table}
\caption{Recorded HTTP headers}
 \begin{tabular}{ | l |  p{3cm} | p{5cm} |}
    \hline
    Header & Reference & Description \\ \hline
\httph{Age} & RFC7234 & When obtaining response from a cache, the number of
seconds since validation \\ \hline
\httph{Cache-Control} & RFC7234 & Header used for a variety of directives \\ \hline
\httph{Expires} & RFC7234  & Gives the date/time after which the
   response is considered stale. \\ \hline
\httph{Pragma} & RFC7234 & Archaic HTTP 1.0 header  \\ \hline
\httph{Warning} & RFC7234  & For additional information about possible incorrectness \\ \hline
\httph{Content-Type} & RFC7231 & To select the correct parser \\ \hline
\httph{If-None-Match} & RFC7232  & Request header to check if
                                   \httph{ETag} has changed   \\ \hline
\httph{If-Modified-Since} & RFC7232  & Request header to check if
                                   \httph{Last-Modified} has changed    \\ \hline
\httph{Last-Modified} & RFC7232 & When the resource was last modified \\ \hline
\httph{ETag} & RFC7232 & An opaque validator to check if the resource
has changed  \\ \hline
\httph{X-Cache} &  & Inserted by some caches to indicate cache status \\ \hline
\httph{Date} & RFC7231 & The time of the message. Used in conditional requests
and heuristics \\ \hline
\httph{Surrogates} & Edge Architecture Specification\todo{ref} & Draft
to allow more fine-grained control for proxies. \\ \hline
\httph{Client-Aborted} & libwww\todo{ref}  & Header inserted by User
Agent to indicate that it aborted the download \\ \hline
\httph{Client-Warning} & libwww\todo{ref}  & Header inserted by User
Agent to give details about problems with the download \\ \hline

@@

Finally, if the LOV SPARQL endpoint used a URI for the vocabulary that
was different from the namespace URI (after a normalization step),
another request would be made to record the selected HTTP headers from
that as well.
\end{appendices}

>>>>

\subsection{Other caching headers}

The two most important headers to control caching are \httph{Expires}
and \httph{Cache-Control}. As shown in Table~\ref{tab:headers}, there
are also others. We found \httph{Pragma} in 287 responses, but except
for two hosts, where they were superflous, they were only used to
prohibit caching. \httph{Surrogates} were not observed.

@@

We next examine the freshness lifetime of the 2965 information
resources that returned a successful response, either with SPARQL
results or RDF data. 405 returned valid headers, but 114 did so to
\emph{prohibit} caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occured, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching.

Caching may also be prohibited by setting a non-positive freshness
lifetime. Common practice is to set the \httph{Expires} header to a
date far back in time, which has also been done in large parts of the
present dataset. For the following, we have set any non-positive
lifetime to 0, as that is semantically equivalent. With that, we
obtain the summary statistics in Table~\ref{tab:summaryhard}. In
figure\todo{add figure}, we have included a histogram with the counts
of the resources that had 0 lifetime, and then boxes of varying length.

@@

In Figure\todo{figure with over freshness}, there is a barplot where
the freshness lifetime is grouped in bins, where the first bin are the
cases where caching is explicitly prohibited. The next bins are for
lifetimes, where the values are grouped if they are on the order of
seconds, minutes, hours, etc, i.e. the second bin counts the lifetimes
in the interval \[1,59\] seconds, etc. We see that in these
categories, the most common is to prohibit caching. Nevertheless, many
also declare a standards compliant freshness lifetime in minutes to
days.

In Figure\todo{figure with freshness breakdown}, we have broken this
up by the type of resource that was accessed, i.e. SPARQL endpoints,
vocabularies, dataset descriptions or unclassified information
resources. First, we note that it seems like the distribution of
freshness lifetime is quite different for the different types, an
observation that is also supported by a similar hypothesis test as
above, with a \pvalue = 0.00001. We also note that it is often
prohibited to cache dataset descriptions. This is odd, since
statistics about datasets is usually costly to compute and should be
cached. The VoID specification\todo{ref} also notes that the
statistics are considered estimates.

We also note that prohibition of SPARQL results caching is rare,
amongst the servers that expose caching headers, it is common
that the result may be cached for some minutes.\todo{more to point out?}

@@

\begin{table}
\caption{Recorded HTTP headers}\label{tab:headers}
 \begin{tabular}{ | l |  p{3cm} | p{5cm} |}
    \hline
    Header & Reference & Description \\ \hline
\httph{Age} & RFC7234 & When obtaining response from a cache, the number of
seconds since validation \\ \hline
\httph{Cache-Control} & RFC7234 & Header used for a variety of directives \\ \hline
\httph{Expires} & RFC7234  & Gives the date/time after which the
   response is considered stale. \\ \hline
\httph{Pragma} & RFC7234 & Archaic HTTP 1.0 header  \\ \hline
\httph{Warning} & RFC7234  & For additional information about possible incorrectness \\ \hline
\httph{Content-Type} & RFC7231 & To select the correct parser \\ \hline
\httph{If-None-Match} & RFC7232  & Request header to check if
                                   \httph{ETag} has changed   \\ \hline
\httph{If-Modified-Since} & RFC7232  & Request header to check if
                                   \httph{Last-Modified} has changed    \\ \hline
\httph{Last-Modified} & RFC7232 & When the resource was last modified \\ \hline
\httph{ETag} & RFC7232 & An opaque validator to check if the resource
has changed  \\ \hline
\httph{X-Cache} &  & Inserted by some caches to indicate cache status \\ \hline
\httph{Date} & RFC7231 & The time of the message. Used in conditional requests
and heuristics \\ \hline
\httph{Surrogates} & Edge Architecture Specification\todo{ref} & Draft
to allow more fine-grained control for proxies. \\ \hline
\httph{Client-Aborted} & libwww\todo{ref}  & Header inserted by User
Agent to indicate that it aborted the download \\ \hline
\httph{Client-Warning} & libwww\todo{ref}  & Header inserted by User
Agent to give details about problems with the download \\ \hline

@@

Finally, if the LOV SPARQL endpoint used a URI for the vocabulary that
was different from the namespace URI (after a normalization step),
another request would be made to record the selected HTTP headers from
that as well.\todo{say how many requests failed? It was a lot}
\end{appendices}
